Efficient Selection of Time Series Shapelets

### 2024-04-12 Docs/paper/main.tex
Drafting Background and related work

\section{Introduction}
Time series classification is similar to traditional classification where the
difference is that the order of the values of the time series is crucial.
The order holds discriminatory features for the time series. The literature
is rich of algorithms for classification, and one way to group these
algorithms is based on the techniques that each algorithm uses to exploit the
discriminatory features suggested by Bagnall et al.~\cite{bake_off}:
\begin{itemize}
	\item Distance based: Classification is based on using distance measures to
	      the raw time series.
	\item Feature based: Global features are extracted and the data is transformed
	      from the time domain to the feature domain spanned by the extracted
	      features, and the classification is done using a standard classifier.
	\item Interval based: Techniques in this family extract one or more intervals,
	      then statistical summaries of these intervals are used as features.
	\item Shapelet based: This technique exploits discriminatory patterns that
	      appear in time series of the same class.
	\item Dictionary based: These algorithms use as features the count of
	      repeating patterns.
	\item Convolution based: Features used are extracted from time series using
	      convolutions and pooling operations.
	\item Deep learning: Classification is done by feeding the data to neural
	      networks
	\item Hybrid: Algorithms in this category combine two or more approaches from
	      the previous techniques.
\end{itemize}
Our focus in this paper is on extracting discriminating patterns, i.e. shapelets
that represent time series from each class. The extracted patterns then can be 
used by algorithms from the shapelet based technique in order to execute the
classification task.

@@

Instead of evaluating by giving a score to each candidate independently, our
approach use efficient similarity search indexing approaches and exploits the
similarity structure in order to select the best candidates. The idea is that
good shapelets representing a class $C$ should have a lot nearest neighbors
that are from the same class, and at the same time, should have few or no
neighbors from other classes.

@@

\section{Related work}
This~\cite{keogh_shapelet} is where it all started, it has to be here.
This paper~\cite{random_shapelets} uses random shapelet extraction.
This paper~\cite{fss} uses neat tricks select shapelets.
This paper~\cite{learning_shapelets} takes an alternative approach, where 
instead of selecting shapelets it \emph{learns} non existing shapelets.
We should mention the evaluation measures~\cite{alternative_measures}.
We will use Silhouette~\cite{silhouettes} as motivation to exploit neighborhood
structure.

>>>>

\section{Introduction}
Time series classification is similar to traditional classification where the
difference is that the order of the values of the time series is crucial, since 
it holds discriminatory features of the time series. The literature is rich of
algorithms for classifying time series, and one way to group them,
suggested by Bagnall et al.~\cite{bake_off}, is based on the techniques that
each algorithm uses to exploit the discriminatory features:
\begin{itemize}
	\item Distance based: Classification is based on using distance measures to
	      the raw time series.
	\item Feature based: Global features are extracted and the data is transformed
	      from the time domain to the feature domain spanned by the extracted
	      features, and the classification is done using a standard classifier.
	\item Interval based: Techniques in this family extract one or more intervals,
	      then statistical summaries of these intervals are used as features.
	\item Shapelet based: This technique exploits discriminatory patterns that
	      appear in time series of the same class.
	\item Dictionary based: These algorithms use as features the count of
	      repeating patterns.
	\item Convolution based: Features used are extracted from time series using
	      convolutions and pooling operations.
	\item Deep learning: Classification is done by feeding the data to neural
	      networks
	\item Hybrid: Algorithms in this category combine two or more approaches from
	      the previous techniques.
\end{itemize}
Our focus in this paper is on extracting discriminating patterns, i.e. shapelets
that represent time series from each class. The extracted patterns then can be 
used by algorithms from the shapelet based technique in order to execute the
classification task.

@@

Instead of evaluating by giving a score to each candidate independently, our
approach use efficient similarity search indexing approaches and exploits the
similarity structure in order to select the best candidates. The idea is that
good shapelets representing a class $C$ should have a lot of nearest neighbors
that are from the same class, and at the same time, should have few or no
neighbors from other classes.

\section{Background \& Related work}
In a classification setup, the idea of a shapelet, introduced by Ye and 
Keogh~\cite{keogh_shapelet}, is that time series from the same class must have 
some shared pattern that is similar to all of them and less similar to time 
series from other classes. A pattern is simply a subsequence of the time series,
and the (dis-)similarity is a distance measure, usually the normalized Euclidean
distance. Besides the high accuracy of shapelets, they are also interpretable,
since the subsequences chosen as shapelets can be interpreted as the event that
happened at that period of time.
Although the idea being simple, it introduced many challenges. The two 
challenges we are concerned about are:
\begin{enumerate}
  \item How to choose patterns?
  \item How to evaluate them?
\end{enumerate}
In the original work~\cite{keogh_shapelet}, the authors considered all the 
subsequences as patterns, and for the evaluation part, they ranked the patterns
by assigning to each a score. The score was the information gain, which is based
on the Shanon entropy, before and after considering the subsequence. At the end,
only the subsequences with the highest score were selected as shapelets.
Despite that only few patterns were selected, they all had to be evaluated,
which means, the similarity between all of them had to be computed in order to 
assign a score. Computing the similarity was the bottleneck of the computation,
in order to speed it up, the authors had to apply early computation pruning.
The pruning resulted in a speed up, but still the approach is very slow,
and not practical for medium and large size datasets. To improve the running 
time, random selection of shapelets~\cite{random_shapelets} have been 
introduced. The authors empirically shown that by selecting a large number of
shapelets, i.e., by randomly selecting subsequences with different lengths from 
random time series at random positions, the computation becomes much faster
since only few patterns are considered, while the accuracy remains competitive.

In order to speed up the selection process, Ji et al.~\cite{fss} introduced 
Fast Shapelet Transform (FSS), that works on two different steps. First, they 
sample  time series with the help of subclass splitting. This reduces the 
number of time series to be considered significantly, without affecting the 
accuracy, since the goal is to look for similar patterns and not all patterns.
Second, the Local Farthest Deviation Points (LFDP) are selected and used as the
endpoints of the subsequences. An LFDP is the point that has the largest 
distance from the subsequence fitting line. The intuition is that the constant
subsequences should be ignored and only subsequences where there are changes
should be interesting.

Regardless of the technique used in order to select the candidates, the 
evaluation in the above approaches was done using the information gain.
Lines et al.~\cite{alternative_measures} were the first to consider using
different evaluation methods instead of the information gain. They considered
as evaluation measures statistical tests to verify if points from the same
distribution or not, such as Kruskal Wallis and Mood's medians. On another 
the same authors use the ANOVA F-statistic~\cite{shapelet_transform} as an 
evaluation measure, which was not significantly better than the information 
gain, but on average, shapelets selected using the F-statistic resulted in high
classification accuracy.

This paper~\cite{learning_shapelets} takes an alternative approach, where 
instead of selecting shapelets it \emph{learns} non existing shapelets.
We should mention the evaluation measures~\cite{alternative_measures}.
We will use Silhouette~\cite{silhouettes} as motivation to exploit neighborhood
structure.

@@

\section{Selection}
